2.4 PRNT50

### 2025-05-12 2_scripts/2.4 PRNT50.qmd
Add Panel B for Figure S.2

# APL Data
load(file = "../1_data/private/APLs.RData") # a clean subset
```

@@

```{r}
APLs<-merge(APLs, select(RFD4682_ev2, order_ID, 
                   `RBD-IgGII_RESULT`, `RBD-IgGII_INTERP`,`RBD-IgGII_RESULT_DATE`
                   ), by="order_ID",all=T)
colnames(APLs)<-c(names(APLs)[1:10], "abbott_s", "pos_abbott_s", "month_abbott_s")

@@

```{r}
CBSs$pos_roche_s<-CBS$interp_roche_s
CBSs_AB<-CBSs[CBSs$province=="AB",]

@@

@@

#### 2.2 Roche-S \<--\> Abbott-S transformation

@@

The APL data Carmen shared with us has vaccination info, the 2024 Jan ver of it even has the vax brand. While the CBS data we have doesnot include vax info. However, a paper written by Sheila has the vax% of those in CBS data, which we can use for comparison on vax%.

@@

```{r}
# create vax% cat var
APLs$vax_cat<-rep(NA, length(APLs$clean_IDe))
APLs$vax_cat[APLs$vax %in% c("1 Dose","1-2 Doses","2 Doses","3 Doses",
                             "4 Doses")]<-1
APLs$vax_cat[APLs$vax %in% c("Not Immunized", "Pre-Immunization")]<-0
table(APLs$vax, APLs$vax_cat, useNA = 'ifany')
```

@@

APLs$vax_est<-rep(0, length(APLs$clean_IDe))
APLs$vax_est[APLs$pos_abbott==0 & APLs$pos_abbott_s==1]<-1
APLs$vax_est[is.na(APLs$pos_abbott) | is.na(APLs$pos_abbott_s)]<-NA
table(APLs$vax_est)
```

@@

```{r}
p12<-dta.vax %>%
  ggplot(aes(x=samplemonth, y=prop.vax, group=data, colour=data)) +
  geom_point(aes(colour=data)) +
  geom_line(aes(colour=data))  +
  theme_bw() + 
  ylim(0,1)+
  ylab("Vaccinated%") +
  scale_colour_manual("", values = c("orange", "red", "red", "purple"),
                      labels = c("AB General Pop.", "APL Vax%", 
                                 "Anti-S(+)", "Anti-S(+) N(-)")
                      )
#  scale_linetype_manual("", values=c(3,3,1,1), labels = p09.lab) 
p12
```

#### 4.2.4 Apply (S+ N-)/N- for vax% estimation

@@

p12.lab <- c("AB General Pop.", "APL Vax%", 
             "APL: Anti-S(+)", "APL: Anti-S(+) N(-)", "CBS: Anti-S(+) N(-)")
p12<-dta.vax %>%
  ggplot(aes(x=samplemonth, y=prop.vax, colour=interaction(study, adj),
                                        linetype=interaction(study, adj))) +
  geom_point() +
  geom_line()  +
  theme_bw() + 
  ylim(0,1)+
  ylab("Vaccinated%") +
  scale_colour_manual("", values = c("orange", "red", "red","purple", "blue"),
                      labels = p12.lab)+
  scale_linetype_manual("", values=c(1,1,3,3,3), labels = p12.lab) 
p12
```
#### Print p12
```{r}
png(file="../4_output/figs/Fig.S2 Vax.png", 
    width=2400, height=1440, res=300)
p12
dev.off()
```

@@

RFD4682_ev2 %>%
  filter(NAAT_STATUS == "Positive") %>%
  group_by(NAAT_month_cat) %>%
  count(`N-IgG_INTERP`)
```

>>>>

# APL Data
load(file = "../1_data/private/APLs.RData") # a clean subset
# Updated APL v2 data (ID cleaned and encrypted)
load(file = "../1_data/private/RFD4682_ev2.RData")
```

@@

```{r}
APLs<-merge(APLs, select(RFD4682_ev2, order_ID, 
                   `RBD-IgGII_RESULT`, `RBD-IgGII_INTERP`,
                   `RBD-IgGII_RESULT_DATE`,
                   VACCINATION_STATUS,
                   ), by="order_ID",all.x=T, all.y=F)

# Remove duplicated records (N=4)
APLs<-APLs %>%
  filter(order_ID %in% RFD4682_ev2$order_ID) %>%
  rename(abbott_s = `RBD-IgGII_RESULT`,
         pos_abbott_s = `RBD-IgGII_INTERP`,
         date_abbott_s = `RBD-IgGII_RESULT_DATE`,
         vax = VACCINATION_STATUS)
APLs$month_abbott_s<-as.yearmon(APLs$date_abbott_s)

@@

```{r}
CBS<-arrange(CBS, pid)
CBSs<-arrange(CBSs, pid)
CBSs$pos_roche_s<-CBS$interp_roche_s
CBSs_AB<-CBSs[CBSs$province=="AB",]

@@

```{r}
# Save the modified CBSs, APLs, and CBSs_AB data
save(APLs, file = "../1_data/private/APLs.RData")
save(CBSs, file = "../1_data/private/CBSs.RData")
save(CBSs_AB, file = "../1_data/private/CBSs_AB.RData")
```

@@

#### 2.2 Roche-S --> Abbott-S transformation
This part is skipped and not mentioned in the manuscript.

@@

The APL data Carmen shared with us has vaccination info, the 2024 Jan ver of it even has the vax brand. While the CBS data we have no vax info. However, a paper written by Sheila has the vax% of those in CBS data, which we can use for comparison on vax%.

@@

```{r}
# create vax% cat var
APLs$vax_cat<-rep(NA, length(APLs$order_ID))
APLs$vax_cat[APLs$vax %in% c("1 Dose","1-2 Doses","2 Doses","3 Doses",
                             "4 Doses")]<-1
APLs$vax_cat[APLs$vax %in% c("Not Immunized", "Pre-Immunization")]<-0
table(APLs$vax, APLs$vax_cat, useNA = 'ifany')
```

@@

APLs$vax_est<-rep(0, length(APLs$order_ID))
APLs$vax_est[APLs$pos_abbott==0 & APLs$pos_abbott_s==1]<-1
APLs$vax_est[is.na(APLs$pos_abbott) | is.na(APLs$pos_abbott_s)]<-NA
table(APLs$vax_est)
```

@@

```{r}
p12<-dta.vax %>%
  ggplot(aes(x=samplemonth, y=prop.vax, group=data, colour=data)) +
  geom_point(aes(colour=data)) +
  geom_line(aes(colour=data))  +
  theme_bw() + 
  ylim(0,1)+
  ylab("Vaccinated%") +
  scale_colour_manual("", values = c("orange", "red", "red", "purple"),
                      labels = c("AB General Pop.", "APL Vax%", 
                                 "Anti-S(+)", "Anti-S(+) N(-)")
                      )
#  scale_linetype_manual("", values=c(3,3,1,1), labels = p09.lab) 
p12
```
#### 4.2.4 Apply (S+ N-)/N- for vax% estimation

@@

p12.lab <- c("AB General Pop.", "APL Vax%", 
             "APL: Anti-S(+)", "APL: Anti-S(+) N(-)", "CBS: Anti-S(+) N(-)")
p12<-dta.vax %>%
  ggplot(aes(x=samplemonth, y=prop.vax, colour=interaction(study, adj),
                                        linetype=interaction(study, adj))) +
  geom_point() +
  geom_line()  +
  theme_bw() + 
  ylim(0,1)+
  ylab("Vaccinated%") +
  scale_colour_manual("", values = c("orange", "red", "red","purple", "blue"),
                      labels = p12.lab)+
  scale_linetype_manual("", values=c(1,1,3,3,3), labels = p12.lab) +
  theme(legend.position = c(0.8,0.25))
p12
```
#### Print p12
```{r}
png(file="../4_output/figs/Fig.S2 Vax.png", 
    width=2400, height=1440, res=300)
p12
dev.off()
```
#### 4.2.5 Figure S2 Panel B using transformed data
```{r}
# APL S+/N- N is transformed (Abbott-->Roche)

table(APLs$pos_abbott, APLs$pos_roche, useNA = 'ifany')
# if transformed into Roche, N=11501 participants will be identified as positive

table(APLs$pos_abbott_s, APLs$pos_roche, useNA = 'ifany')

# Add vax indicator using S+ and N- (transformed N Abb->Ro)
APLs$vax_est_t<-rep(0, length(APLs$order_ID))
APLs$vax_est_t[APLs$pos_roche==0 & APLs$pos_abbott_s==1]<-1
APLs$vax_est_t[is.na(APLs$pos_roche) | is.na(APLs$pos_abbott_s)]<-NA
table(APLs$vax_est_t, useNA = 'ifany')

tbNT_vax_est<-APLs %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            n_pos = sum(pos_roche, na.rm = T),
            s_pos = sum(pos_abbott_s, na.rm = T),    
            n_vax= sum(vax_est_t, na.rm = T)
            )
tbNT_vax_est$n_neg = tbNT_vax_est$n - tbNT_vax_est$n_pos
tbNT_vax_est$prop.vax<-tbNT_vax_est$n_vax/tbNT_vax_est$n_neg

# clean the table
tbNT_vax_est<-tbNT_vax_est[!is.na(tbNT_vax_est$samplemonth),]
tbNT_vax_est$prop.vax[tbNT_vax_est$prop.vax==0]<-NA

## add CBS data to merged data
d06<-select(tbNT_vax_est, samplemonth, prop.vax)
d06$data<-rep("S_pos_N_Neg (Anti-N Transformed)", length(d06$samplemonth))

#dta.vax<-rbind.data.frame(dta.vax, d05)
```

```{r}
# CBS S+/N- N is transformed (Roche-->Abbott)

# Add vax indicator using S+ and N- (transformed N Abb->Ro)
CBSs_AB$vax_est_t<-rep(0, length(CBSs_AB$pid))
CBSs_AB$vax_est_t[CBSs_AB$pos_abbott==0 & CBSs_AB$pos_roche_s==1]<-1
CBSs_AB$vax_est_t[is.na(CBSs_AB$pos_abbott) | is.na(CBSs_AB$pos_roche_s)]<-NA
table(CBSs_AB$vax_est_t, useNA = 'ifany')

tbCBST_vax_est<-CBSs_AB %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            n_pos = sum(pos_abbott, na.rm = T),
            s_pos = sum(pos_roche_s, na.rm = T),    
            n_vax= sum(vax_est_t, na.rm = T)
            )
tbCBST_vax_est$n_neg = tbCBST_vax_est$n - tbCBST_vax_est$n_pos
tbCBST_vax_est$prop.vax<-tbCBST_vax_est$n_vax/tbCBST_vax_est$n_neg

## add CBS data to merged data
d07<-select(tbCBST_vax_est, samplemonth, prop.vax)
d07$data<-rep("CBS: S_pos_N_Neg (trasformed)", length(d07$samplemonth))

dta.vax.2<-rbind.data.frame(d06, d07, d02, d05)

```

```{r}
# Use figure to present data
p13.lab <- c("CBS Raw: Anti-S(+) N(-)", "CBS Transformed: Anti-S(+) N(-)", 
             "APL Raw: Anti-S(+) N(-)", "APL Transformed: Anti-S(+) N(-)")
p13<-dta.vax.2 %>%
  ggplot(aes(x=samplemonth, y=prop.vax, colour=data, linetype=data)) +
  geom_point() +
  geom_line()  +
  theme_bw() + 
  ylim(0,1)+
  ylab("Vaccinated%") +
  scale_colour_manual("", values = c("blue", "brown", "purple","green"),
                      labels = p13.lab)+
  scale_linetype_manual("", values=c(3,3,3,3), labels = p13.lab)+
  theme(legend.position = c(0.8,0.2))
p13
```
##### Print p12 and p13 together
```{r}
p1213<-plot_grid(p12 + xlab ("Time of Sample Collection") +
                 coord_cartesian(xlim=as.yearmon(c("Oct 2020", "Feb 2023")))+
                 theme(legend.position = c(0.67,0.25)), 
                 p13 + xlab ("Time of Sample Collection") + ylab("") +
                 coord_cartesian(xlim=as.yearmon(c("Oct 2020", "Feb 2023")))+
                 theme(legend.position = c(0.65,0.2)),
                 labels = c("A", "B"),
                 nrow=1,  align = "h")
p1213

png(file="../4_output/figs/Fig.S2 Vax.png", 
    width=2400, height=1200, res=300)
p1213
dev.off()
```

@@

RFD4682_ev2 %>%
  filter(NAAT_STATUS == "Positive") %>%
  group_by(NAAT_month_cat) %>%
  count(`N-IgG_INTERP`)
```
#### 5.3 Modify the legends and y-axis labels of figure 1
```{r}
# Check the codes in qmd script 2.3
# use key words "Fig.1" "Fig.2" to search
```
